### 2021-05-02 polyphony.tex
lemmas n-best, application, debut intro

\section{Application}
Symbolic Automated Music Transcription
and analysis of music performances

\subsection{Time Scales}
Real-Time Unit (RTU) = seconds

\noindent 
Musical-Time Unit (MTU) = number of measures

\noindent 
conversion via tempo value

\subsection{Representation of Music Performances}
A musical performance is represented symbolically as a finite sequence of timestamped events.
This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1.

\noindent
We consider an infinite alphabet $\Sigma$ of MIDI-like 
events made of:
\begin{itemize}
\item a timestamp in RTU\\ %(real or int = nb samples ?)
      or IOI in RTU
\item a pitch value in 0..127
\item \textsf{on} | \textsf{off} flag
\item a velocity value in 0..127
\end{itemize}