### 2021-05-18 main2.tex
revise V2:  extended expressiveness

% fork of main.tex with more expressive models
% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
% SWA read tape (under head) at every transition (even epsilon)
% SW-VPA read top of stack (when not empty) at every transition

\documentclass[runningheads]{llncs}
%\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{hyperref}
%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
%usepackage{url}

% for footnote ref
\usepackage{refcount}

% array and tabular
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

% extension of enumerate env. (style for displaying counters)
% \usepackage{enumerate}

%% pictures
% \usepackage{graphicx} 
% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
% \graphicspath{fig/}

%% PGF, Tikz
\usepackage{tikz-cd}
%% \usepackage{pgfplots}
%% \usepgfplotslibrary{dateplot}
%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
% \usepackage{tikz}
% \usetikzlibrary{cd}
%% \usetikzlibrary{arrows}
%% \usetikzlibrary{calc}
%% \usetikzlibrary{snakes}
%% \usetikzlibrary{backgrounds}
% \usetikzlibrary{trees}
%% \usetikzlibrary{automata}
%% \usetikzlibrary{positioning}
%% \usetikzlibrary{matrix}
%% \usetikzlibrary{patterns}
%% \usetikzlibrary{shapes}

% symbols
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsbsy}
\usepackage{bbold}
\usepackage{latexsym}
%\usepackage{amsfonts}
\usepackage{stmaryrd}
%\usepackage{mathabx}
%\usepackage{MnSymbol}
\usepackage{harmony} % simple music fonts
\usepackage{mathtools} % for arrows
%\usepackage{mathptmx}

%% theorem environments
\usepackage{theorem}
% \newtheorem{theorem}{Theorem} %[section]
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{proposition}[theorem]{Proposition}
% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
% \theorembodyfont{\slshape}
% \newtheorem{example}[theorem]{Example}
% \newtheorem{remark}[theorem]{Remark}
\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

%% algorithms
%\usepackage{algorithm} 
%\usepackage{program}

%% for new macros
\usepackage{xspace}

%% arrows etc
%\input{rewriting}

%% music symbols
% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
\usepackage{musicography}

%% Misc macros

\def\ie{\textit{i.e.}\xspace}
\def\eg{\textit{e.g.}\xspace}
\def\wrt{\textit{wrt}\xspace}
%\def\wlog{\textit{wlog}\xspace}
\def\etc{\textit{etc}\xspace}

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\T}{\mathcal{T}}

\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}

%\sloppy

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
%\title{Symbolic Weighted Parsing and Automated Music Transcription}
%\title{Symbolic Weighted Language Models and Automated Music Transcription}
\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
\author{Florent Jacquemard}
\institute{INRIA \& CNAM, Paris, France\\
\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
%\authorrunning{Florent Jacquemard}

\date{\today}

\begin{document}
\thispagestyle{empty}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro}
\input{introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SWT & SWA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Automata and Transducers}
\label{section:transducer}\label{sec:transducer}
\label{section:SWA}\label{sec:SWA}
\label{section:SWT}\label{sec:SWT}

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
between words and languages, using weighted transducers, 
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers. 
%
The models introduced in this section generalize 
weighted automata and transducers~\cite{Droste09handbook} 
%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labelling each transition with a weight function (instead of a simple weight value), 
that takes the input and output symbols as parameters. 
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an generic semiring, 
whereas the latter guards are restricted to the Boolean semiring.

\subsection{Semirings} 
\label{section:semiring}
\input{semiring}

\subsection{Label Theory}
\label{section:symbols}
\input{labels}

\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)

Let $\Semiring$ be a commutative and complete semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a 4-uplet of transition functions 
$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
\end{definition}

A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
is called an \emph{$\epsilon$-transition}.
%
For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
we shall sometimes write, overloading the function names:
$\wei_{ij}(q, a, b, q') = \eta(a, b)$ when  $a \in \Sigma$,  $b \in \Delta$.

\noindent 
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$, resp. $t$, being called input word, resp. output word.
%
% \noindent
More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
\[
\begin{array}{lccl}
\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
\weight_T(q, au, bv, q') & = & 
   \displaystyle\bigoplus_{q'' \in Q} &
   \wei_{00}(q, a, b, q'') \otimes \weight_T(q'', au, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{10}(q, a, b, q'') \otimes \weight_T(q'', u, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{01}(q, a, b, q'') \otimes \weight_T(q'', au, v, q')\\
  & \oplus & \displaystyle\bigoplus_{q'' \in Q} &
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\
\end{array}
\]
%
We recall that, by convention, an empty sum with $\bigoplus$ is $\zero$. 
%
The above sum may be infinite because of the $\epsilon$-transitions, % in the first and second lines.
in this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
%
%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***

%** The application of $\oplus$ can be seen as a ND choice between transitions **
% choose to advance head in input, in output or both (or none)
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
read $a$, resp. $b$, at the current position in the input word, resp. output word;
if $i = 1$, resp. $j = 1$,
increment the current position in the input word (otherwise, do not change it);
change state from $q$ to $q'$.
The input and output symbols are always read at current positions, 
even during transition that do not increment the head's position.
This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
which ignore the symbol at current position in this case.
In our case, symbols in a infinite alphabet cannot be stored in the finite memory of the transducer, 
whereas considering the current symbol may be useful to compute a transition weight .
%(even when it does not change the head's position, like with $\epsilon$-transitions).
%
The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
are missing in the definition of $\weight_T$.
It means that $T$ must avoid configurations where it reached the end of 
the output word and not of the input one, or vice-versa.
This can be done by using $\wei_{10}$ and $\wei_{01}$
before reaching the end of word, and using a special state for this purpose.

Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and writing $b$ 
(including the case $a = \epsilon$ and $b = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to a transition's guard not satisfied by $\<a, b>$ in 
the case of symbolic transducers.

The expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei_{10}(q, a, b, q'')$, in the fourth line, is the cost of 
the deletion of the symbol~$a \in \Sigma$ in~$s$, 
$\wei{01}(q, a, b, q'')$, in the fifth line, is the cost 
of the insertion of~$b \in \Delta$ in $t$, 
and $\wei_{11}(q, a, b, q'')$, in the last line, is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

\medskip\noindent
The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
is defined as follows: 
\begin{equation}
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
\label{eq:weightT}
\end{equation}

\begin{example}
We can define with a \SWT the computation of a similarity measure 
between timed sequences
similar to dynamic time warping (DTW).

Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
We consider a \SWT with one state $q$ and transitions
$\wei_{11}(q, d, d', q) = 
 \wei_{01}(q, d, d', q) = 
 \wei_{10}(q, d, d', q) = |d' - d|$,
and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
% needs reading input/output symbols even by epsilon-transitions.
The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
for the computation of an optimal match between words, 
the matching cost for two symbols being the 
the time distance between them.
\endex
\end{example}

\begin{example}
% simple pointwise distance between two sequences of timestamped events **
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
This example of $\Delta$ is motivated by the case of music notation, 
where several notated events (notes) can be tied together, 
with a \emph{tie} or a \emph{dot}
(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
 or equivalently~\musQuarterDotted),
meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \one$, $\init(q_1) = \zero$, 
$\final(q_0) = \one$, $\final(q_1) = \zero$, 
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a pair of transition functions 
$\wei_{0}$, $\wei_{1}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
is called an \emph{$\epsilon$-transition}.

\noindent
As above in the case of \SWT, 
the transition functions $\wei_0$ and $\wei_1$
can be synthesized into a single function
$\wei: Q \times \Sigma \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

\subsection{Properties}
The two following properties will be useful to our approach on 
symbolic weighted parsing in Section~\ref{sec:parsing}.

\begin{proposition}
Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
let 
$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
be the synthesized form of 
$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
Its state entering weight function is defined by 
$\init'(1, q) = \init(q)$ for all $q \in Q$, 
and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and its state leaving weight function by 
$\final'(n, q) = \final(q)$ or all $q \in Q$, 
and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcr}
\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
%
\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
%
\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1,\\ 
%
\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
%
\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
%
\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1.
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_1$.

\noindent
The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
\qed
\end{proof}

The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
where the size $\| T \|$ of $T$ is its number of states $|Q|$.

% gen. to transducers
\begin{proposition}
When $\Semiring$ is commutative, bounded and complete, 
given a \SWA $A$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$
there exists an effectively constructible \SWA $A'$ 
without $\epsilon$-transitions 
over $\Sigma$, $\Semiring$ and $\bar\Phi$,
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof} Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
The automaton $A'$ is $\< Q, \init, \<\wei'_0, \wei'_1>, \final >$,
where, for all $q, q' \in Q$,
$\wei'_0(q, q') = \zero$  
and 
** Tb revised **   
% \[
% \begin{array}{rcccl}
% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
% \end{array}
% \]
\[
\wei'_\Sigma(q, q') = 
\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
\]
In the above definition of $\wei'_\Sigma$ we use the operator
of product of function of $\Phi_\Sigma$ by $\Semiring$ 
described in Section~\ref{section:symbols}.
%
By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
** NO. TBC see \cite{Mohri02ijfcs} **
it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
\qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SW-VPA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The following model generalizes Symbolic VPA~\cite{dAntonyAlur14SVPDA}
from Boolean semirings to arbitrary semiring weight domains.

\subsection{Definition}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Omegai$, $\Omegac$, $\Omegac$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and  
let  $(\Phi_\epsilon, \Phii, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omegar}$.
%
%Moreover, we extend this theory with a set $\Phii$ 
%of unary functions in $\Omegai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ is a tuple
$A = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
where $Q$ is a finite set of states, 
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and 
$\weii : Q \times Q \to \Phii$,  
$\weic : Q \times Q \times P \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weie : Q \times Q \to \Phir$,  
are transition functions.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Omegai$, 
$\call{c} \in \Omegac$, 
$\return{r} \in \Omegar$, 
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Omegai \times Q \to \Semiring & 
\weii(q, a, q') = \phi_\mathsf{i}(a) & 
\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
%
\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
%
\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
%
\weie: Q \times \Omegar \times Q \to \Semiring & 
\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
\end{array}      
\]

\noindent
The intuition is the following for the above transitions.
\begin{description}
\item $\weii$ : read the input internal symbol $a$, change state to $q'$ (stack is untouched).
\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
\item $\weir$ : when the stack is not empty, 
      read and pop from stack a pair made of $\call{c}$ and $p$, 
      read the input return symbol $\return{r}$, change state to $q'$.
      In this case, the weight function $\phi_\mathsf{r}$ 
      computes a value of matching between the call and return symbols $c$ and $r$.
      This value might be set to $\zero$ in order to express that the symbols do not match.
\item $\weie$ : when the stack is empty, 
      read the input symbol $\call{r}$, change state to $q'$.
\end{description}

We give now a formal definition of these transitions of the automaton~$A$
in term of %a weight value computed by 
an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
Therefore, $\weight_A$ is a function from 
$Q \times \Gamma^* \times \Omega^* \times Q \times \Gamma^*$ into $\Semiring$.
\[
\begin{array}{rcl}
\weight_A\bigl(\config{q}{\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
%
\weight_A\bigl(\config{q}{\gamma}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, q'', p\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
\end{array}
\]
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
and $\gamma$ is the rest of stack.

\noindent
The weight associated by $A$ to $s \in \Omega^*$
is then defined as follows,
following empty stack semantics: 
\begin{equation}
A(s)  = 
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes} 
\weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr) 
\mathop{\otimes} \mathsf{out}(q').
\label{eq:weightA}
\end{equation}

\begin{example}
structured words...
intro language of music notation ?
\end{example}

\subsection{Properties}
Like VPA and symbolic VPA, 
the class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}
Let $A_1$ and $A_2$ be two (\SWVPA)
over the same $\Omega$ and $\Semiring$.
There exists two $\SWVPA$ $A_1 \oplus A_2$ and $A_1 \otimes A_2$, 
effectively constructible, 
such that for all $s \in \Omega^*$, 
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and 
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
The construction is essentially the same 
as in the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.

\subsection{Best-first Search} 
\label{sec:best}\label{sec:search}
**hypotheses**
Let us assume that the semiring~$\Semiring$ is
commutative, bounded, complete, and total.
and assume an effective label theory.
%s.t. for all $\phi \in \Phii$, 
%$\psi \in \Phir$, 
%and $\eta \in \Phicr$, 
%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
%are computable...

% total ? 
% monotonic and superior writ natural ordering
%Regarding the infinite sum operator, note that
%$\bigoplus_{x \in \Phi_\Omega} \phi(x)$, 
%$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and  
%... exist and in $\Semiring$.

We propose a Dijkstra algorithm computing the minimal weight by $A$, 
\wrt~$\leq_\oplus$, for a word in~$\Omega^*$.

\noindent
More precisely, 
let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
and for every two states $q, q' \in Q$ and $\sigma \in \{ \bot, \top \}$, 
let 
\[ 
  d_0(q, \sigma, q') = \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\config{q}{\sigma}, s, \config{q'}{\sigma}\bigr). 
\]
Since $\Semiring$ is complete, this infinite sum is well defined, 
and since $\leq_\oplus$ is assumed total, it is the minimum in $\Omega^*$
of $s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$ \wrt this ordering.
%
When $\sigma = \bot$, $d_0(q, \sigma, q')$ is the central 
expression in a term of the definition (\ref{eq:weightA}) of $A(s_0)$ for the minimum $s_0$
(for the above function).
When $\sigma = \bot$, intuitively, it is the minimum weight of a computation of $A$
starting in state $q$ with a stack $\gamma \in \Gamma^*$ (possibly empty), 
and ending in state $q$ with the same stack $\gamma$, such that moreover 
the computation does not touch a symbol of $\gamma$. 
That means that during the computation, 
$A$ may apply the first case of in the definition of $\weight_A$ (internal symbol),
as well as the second case, to can push call symbols on the top of $\gamma$, 
and may pop these symbols with the third case (return symbol). 
However, it cannot apply one of the two last cases (return symbol and empty stack)
when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

The algorithm~\ref{algo:Dijkstra}
constructs iteratively a marking $d : Q \times \{ \bot, \top \} \times Q \to \Semiring$
%of the triplets $\<q, \sigma, q'>$ 
%of states of $A$ by weight values in $\Semiring$, 
that converges eventually to $d_0(q, \sigma, q')$. 
%It uses for that purpose a priority queue $P$ containing triplets of 
%$Q \times \{ \bot, \top \} \times Q$.

\begin{algo}[1-best for \SWVPA] \label{algo:Dijkstra}
\textbf{initially} let $P = Q \times \{ \bot, \top \} \times Q$, %contains all %triplets 
%$\< q_1, \bot, q_2>$ and $\<q_1, \top, q_2>$ for $q_1, q_2 \in Q$, with 
and 
$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \one$ if $q_1 = q_2$ and
$d(q_1, \bot, q_2) = d(q_1, \top, q_2) = \zero$ otherwise.

\smallskip\noindent
\textbf{while} $P$ is not empty

\noindent
\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
 such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.

\noindent\quad \textbf{for all} $q_0, q_3 \in Q$ and $p \in P$ do

\noindent
\qquad
\(\begin{array}{rrcl}
 & d(q_1, \sigma, q_3) & \opluseq &
   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, a, q_3)\\
%
%  & d(q_0, \sigma, q_2) & \opluseq & 
%    \displaystyle\bigoplus_{a \in \Omegai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
%
\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})\\
& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
%
\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
  d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Omegar} \weie(q_2, {\return{r}}, q_3)\\
%
%& d(q_0, \bot, q_2) & \opluseq & 
%d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
%
%\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
%d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
%
& d(q_1, \bot, q_3) & \opluseq & 
   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
%
\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
\end{array}
\) 
\end{algo}

\noindent
The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
since~$\Semiring$ is complete.
The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
and each iteration has a time complexity $O(|Q|^2 . |P|)$.
This gives a time complexity $O(|Q|^4 . |P|)$. 
It can be reduced by implementing $P$ as a priority queue, 
prioritized by the value returned by $d$
***complete***. %$|Q|^3.\log(|Q|^2)$

The correctness of Algorithm~\ref{algo:Dijkstra} 
is ensured by the invariant expressed in the following lemma.
\begin{lemma}
For all $\< q_1, \sigma, q_2> \notin P$, $d(q_1, \sigma, q_2) =  d_0(q_1, \sigma, q_2)$.
\end{lemma}
The proof is by contradiction, 
assuming a counter-example minimal in the length of the witness word.

% \begin{algorithm}
% \caption{<your caption for this algorithm>}
% \label{<your label for references later in your document>}
% \begin{program}
% \seq{|line| := `` '', i := 1};
% \WHILE i \neq n+1 \DO
%       |line| := |item|[i] \concat `` '' \concat |number|[i]; 
%       i := i+1; 
%       \WHILE i \neq n+1 \AND |item|[i] = |item|[i-1] \DO 
%       |line| := |line| \concat ``, '' \concat |number|[i]);
%       i := i+1 \OD ; 
%       |write|(|line|) \OD 
% \end{program}
% \end{algorithm}

% The weight of a transition acts as a guard: 
% a transition is activated for a symbol $`a`$ iff its weight $`\phi(a) \neq 0`$ (or $`\psi(a, b) \neq 0`$), 
% the absorbing element.

%\begin{lemma}
\noindent
For computing the minimal weight of a computation of $A$, we use the fact that,
at the termination of Algorithm~\ref{algo:Dijkstra}, %it holds that,
%There exist $q_1, q_2 \in Q$
\[
  {\displaystyle \bigoplus_{s \in \Omega^*} A(s)} = 
  {\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
  \mathsf{in}(q) \mathop{\otimes} d(q, \bot, q') \mathop{\otimes} \mathsf{out}(q').
\]
%\end{lemma}

%\medskip
\noindent
In order to obtain effectively a witness 
(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
we require an additional property the of weight functions.
%
\begin{definition}
Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
A function $\phi$ from $\Omega^n$ into $\Semiring$
is called $k$-\emph{convex} for a natural number $k$ iff 
$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
\end{definition}
A label theory is $k$-convex if all its functions are $k$-convex.

\begin{proposition}
For a \SWVPA $A$ over a 
commutative, idempotent, superior, total and complete semiring 
and an alphabet $\Omega$ with a $k$-convex label theory,
one can construct in PTIME a word $s \in \Omega^*$ 
such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
%$\leg_\oplus$.
\end{proposition}

\subsection{Trees and Nested-Words}
\label{sec:trees}
The hierarchical structure of nested-words, defined the \emph{call} and \emph{return} markup symbols of 
suggest a correspondence between these word and trees. 
The lifting of this correspondence to languages, respective of tree automata and VPA have
been discussed in~\cite{AlurMadhusudan09nested}, 
see also~\cite{Caralp12VPAmult} for the weighted case.
In this section, we discuss the correspondence between the symbolic-weighted extensions
of tree automata and VPA.

Let $\Omega$ be a countable ranked alphabet, such that 
every symbol $a \in \Omega$ has a rank 
$\rank(a) \in [0..M]$ where $M$ is a fixed natural number.
We write $\Omega_k$ the subset of all symbols $a$ of $\Omega$
with $\rank(a) = k$, where $0 \leq k \leq M$.
%$\cup_{i = 0}^{i = M} \Omega_i$ 
%
$\T(\Omega)$ denotes the free $\Omega$-algebra of finite, ordered, 
$\Omega$-labeled trees, 
which is the smallest set such that  $\Omega_0 \subset \T(\Omega)$
and for all $1 \leq k \leq M$, all $a \in \Omega_k$, 
and all $t_1, \ldots, t_k \in \T(\Omega)$, $a(t_1, \ldots, t_k) \in \T(\Omega)$.
%
% tree = single node (leave) labeled with a symbol of $a \in \Omegai$
% (such a tree is simply denoted by $a$)
% or the composition, denoted by $b(t_1,\ldots, t_n$) of a node labeled with $b$
% and $n$ subtrees $t_1$,\ldots, $t_n$.

\renewcommand{\call}[1]{\ensuremath \langle_{#1}}
\renewcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$

Let $\hat\Omega$ be the countable (unranked) alphabet obtained from 
$\Omega$ as follows: 
$\hat\Omega = \< \Omegai, \Omegac, \Omegar >$, with
$\Omegai = \Omega_0$, 
$\Omegac = \{ \; \call{a} \mid a \in \Omega_{>0} \}$,
$\Omegac = \{ \; \return{a} \mid a \in \Omega_{>0} \}$,
where $\Omega_{>0}$ denotes $\bigcup_{k = 1}^{M} \Omega_k$.

\noindent
We define a linearization of trees of $\T(\Omega)$ into 
words of $\hat{\Omega}^*$ as follows:
\begin{description}
\item $\lin(a) = a$ for all $a \in \Omega_0$, 
\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) = 
       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$ 
       when $b \in \Omega_k$, $1 \leq k \leq M$.
\end{description}

% A \emph{regular tree grammar} over $\Omega$ 
% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
% $N$ is a finite set of non-terminal symbols denoted $q$..., 
% $q_\mathsf{i} \in N$ is the starting non-terminal, 
% $R$ is a finite set of production rules of the form
% $q_0 \to a(q_1\ldots q_k)$ where 
% $q_0, q_1, \ldots, q_k \in N$
% $a \in \Omega_k$.
% A tree $t \in \T(\Omega)$ is in the language of $G$ 
% if it can be generated from $q_\mathsf{i}$ by 
% non terminal replacement following the rules of $R$.

Let us assume a label theory $\Phi_{\Omega_k}$ 
for each $k \in [0..M]$.
%
\begin{definition}  \label{def:SWTA}
A \emph{symbolic-weighted tree automaton} (\SWTA)
over the ranked input alphabet~$\Omega$
and the commutative semiring $\Semiring$ 
is a triplet $A = \< Q, \init, \bar{\wei} >$ where
$Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
and $\bar{\wei}$ is a $M+2$-uplet of transition functions made of:
$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
for each $k \in [0..M]$, $\wei_{\Omega, k}$
from $Q \times Q^{k}$ into $\Phi_{\Omega_k}$.
\end{definition}
%
Like in Section~\ref{sec:SWAdef}, we define from $\bar{\wei}$ 
a transition function~$\wei$, %from $\bar\wei$
from 
$Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k$
into~$\Semiring$: %also called $\wei$ for simplicity, 
%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
\[
\begin{array}{rcll}
\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
\quad\mathrm{where~} \phi_{\Omega, k} = \wei_{\Omega, k}(q_0, q_1\ldots q_k).\\
\end{array}      
\]
Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
of a regular tree grammar~\cite{tata}, 
that replaces the non-terminal symbol $q_0$, by $a(q_1, \ldots, q_k)$. 
%in a step of tree building.
%
Such a grammar computes the weights of the derivation trees 
of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.

\noindent
The \SWTA of Definition~\ref{def:SWTA} defines a mapping 
from trees of $\T(\Omega)$ into the weights of~$\Semiring$,
based on the intermediate function $\weight_A$
defined as follows for $q_0 \in Q$ and 
$t = b(t_1,\ldots, t_k) \in \T(\Omega)$,
åwith $0 \leq k \leq M$:
\[
\begin{array}{rccl}
\weight_A(q_0, t) & = & 
   \displaystyle\bigoplus_{q_1 \in Q} &
   \wei(q, \epsilon, q_1) \otimes \weight_A(q_1, t)\\
 & \oplus & \displaystyle\bigoplus_{q_1 \ldots q_k \in Q^k} &
              \wei(q_0, b, q_1 \ldots q_k ) 
   \otimes \displaystyle\bigotimes_{i=1}^{k}
           \weight_A(q_i, t_i).\\
\end{array}
\]

\medskip\noindent
The weight associated by $A$ to  $t \in \T(\Omega)$ is then
\begin{equation}
A(t)  = 
\displaystyle\bigoplus_{q \in Q} \mathsf{in}(q) \mathop{\otimes} \weight_A(q, t).
\label{eq:weightTA}
\end{equation}

\begin{lemma}  \label{lem:SWTA}
For all \SWTA $A$ over~$\Omega$ and $\Semiring$, without $\epsilon$-transitions,
there exists an effectively constructible \SWVPA $A'$ over 
$\hat\Omega$ and $\Semiring$
such that for all $t \in \T(\Omega)$, $A'\bigl(\lin(t)\bigr) = A(t)$.
\end{lemma} 
% 
\begin{proof}
Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is summarized as above by a function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times \bigcup_{k=0}^{M} Q^k \to \Semiring$.

We build 
$A' = \< Q', P', \init', \weii, \weic, \weir, \weie, \final' >$,
computing over $\hat\Omega = \< \Omegai, \Omegac, \Omegar >$,
%
%$\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ and $\Semiring$ 
where $Q' = \bigcup_{k=0}^{M} Q^k$ be  the set of sequences of state symbols of $A$, 
of length at most $M$, including the empty sequence denoted by $\epsilon$, 
and where $P' = Q'$.

\[
\begin{array}{ll}
%\weii: Q \times \Omegai \times Q \to \Semiring & 
\weii(\bar{q}, a, \bar{q} q') = \wei(q', a, \epsilon) & 
\mathrm{for~all~} a \in \Omega_0,\\
%
%\weic: Q \times \Omegac \times Q \times P \to \Semiring & 
\weic(\bar{q}, \call{b}, \epsilon, \bar{q}) = \one & %\wei(p', b, \bar{p}) & 
\mathrm{for~all~} b \in \Omega_{>0},\\
%
%\weir: Q \times \Omegac \times P \times \Sigmar \times Q \to \Semiring & 
\weir(\bar{q}, {\call{b}}, \bar{p}, {\return{b}}, \bar{p} q') = \wei(q', b, \bar{q}) & 
\mathrm{for~all~}  b \in \Omega_{>0},\\
%
%\weie: Q \times \Omegar \times Q \to \Semiring & 
\weie(\bar{p}, {\return{b}}, \bar{q}) = \zero &
\mathrm{for~all~}  b \in \Omega_{>0}.\\
\end{array}      
\]

In practice, it is sufficient to consider in $Q'$ only the prefixes of 
sequences.
\qed\end{proof}

\section{Symbolic Weighted Parsing}
\label{sec:parsing}
Let us now use the models and results of the former sections in %order to define 
an approach to the problem of parsing over infinite alphabet. %appropriate
%
Besides considering infinitely many possible of input symbols, 
handled with suitable language formalisms, 
this approach extends conventional parsing 
by computing a derivation tree modulo 
a generic distance between words,  
defined by a SW transducer given in input.
This enables considering finer word relationships than strict equality 
as in the conventional parsing approach, 
opening possibilities of quantitative analysis via this method.

\subsection{Definition}

Let $\Sigma$ be a countable input alphabet 
and $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$ an output countable alphabet,
let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative, bounded, complete 
and total semiring 
and let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$, 
assumed computable and $k$-convex for some fixed $k$.

\noindent
Assuming given in input:
\begin{itemize}
\item a \SWT $T$ over $\Sigma$,  $\Omega$, and $\Semiring$,
defining a measure %between words 
$T: \Sigma^* \times \Omega^* \to \Semiring$,
\item a \SWVPA $A$ over $\Omega$, and $\Semiring$, defining a series of nested words
      $A : \Omega^* \to \Semiring$,
\item an input word $s \in \Sigma^*$,
\end{itemize}
the problem of \emph{symbolic weighted parsing} is to find 
a nested word $t \in \Omega$ minimizing
\( T(s, t) \otimes A(t)\)
\wrt $\leq_\oplus$, 
\ie such that %Hence, it is the problem of finding 
\( T(s, t) \otimes A(t) = 
  \displaystyle\bigoplus_{t' \in \Omega^*} T(s, t') \otimes A(t') \).

\noindent 
Therefore, it is the problem of optimizing a measure 
called the \emph{edit-distance between $s$ and $A$} in~\cite{Mohri03EDWA}.
%
The input language can also be expressed as a \SWTA, or, 
as a particular case, as a weighted context-free grammar, 
converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
%
In the case of finite alphabets, 
the problem of searching, in a WTA language, 
for the best parse tree for a given,  
sometimes referred as~\emph{weighted parsing}
(see \cite{Goodman99SemiringParsing,MorbitzVogler19weighted-parsing} **more general problems**)
is a particular case of SW parsing.
%
Indeed, it corresponds to the case where $T$
accepts only the pairs $\<s, t>$ such that 
$s$ is the projection of $t$ on $\Omegai$. 
This can be done with a single state $q$ and 
with transition rules of the form:
\begin{description}
\item[] $\wei(q, \epsilon, a, q) = \one$ for all $a \in \Omegac \cup \Omegar$,
\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Omegai$,
\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Omegai$, $a \neq b$.
\end{description}

\subsection{Computation}

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
and the computation time of the functions of the label theory.
\end{proposition}
%
%\begin{proof}
%Bar-Hillel construction
%\qed  
%\end{proof}

%\input{transcription}

\section*{Conclusion}
% discussion

\noindent
-- summary

\noindent
-- other theoretical properties of SW models

\noindent
-- room to improve complexity for best-search algorithm
... modular approach with oracles ...

and extention to $n$-best

\noindent
-- offline algorithm, 
semi-online implementation for AMT (bar-by-bar approach)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BIBLIO                                                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{plain}
%\bibliographystyle{plainurl} 
\bibliographystyle{abbrv}
%\bibliographystyle{splncs04}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% APPENDIX                                                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix

\end{document}

\section{Edit-Distance}

%\subsection{Distance between words or languages}
...algebraic definition of edit-distance of Mohri, in \cite{Mohri03EDWA}
% Mehryar Mohri 
% Edit-distance of weighted automata: General definitions and algorithms
% International Journal of Foundations of Computer Science 14.06 (2003): 957-982.
distance $d$ over $\Sigma^* \times \Sigma^*$ 
into a semiring  $\Semiring = ( \Semiring, \oplus, \zero, \otimes, \one)$.

%\noindent
Let $\Omega = \Sigma \cup \{ \epsilon \} \times \Sigma \cup \{ \epsilon \} \setminus \{ (\epsilon, \epsilon) \}$,
and let $h$ be the morphism from $\Omega^*$ into $\Sigma^* \times \Sigma^*$  
defined over the concatenation of strings of $\Sigma^*$ (that removes the $\epsilon$'s).
%
\noindent
An \emph{alignment} between 2 strings  $s, t \in \Sigma^*$ is an element $\omega \in \Omega^*$ 
such that $h(\omega) = (s, t)$.
%
\noindent
We assume a base cost function $\Omega$ : $\delta: \Omega \to S$, extended to $\Omega^*$ as follows  
(for $\omega \in \Omega^*$): 
\(
\displaystyle\delta(\omega) = \bigotimes_{0 \leq i < |\omega|} \delta(\omega_i)
\).

\noindent
\begin{definition}
For  $s, t \in \Sigma^*$, the edit-distance between $s$ and $t$ is  
\( 
d(s, t) = \displaystyle\bigoplus_{\omega \in \Omega^*\, h(\omega) = (s, t)} \delta(\omega)
\).
\end{definition}

e.g. Levenstein edit-distance: $S$ is min-plus and $\delta(a, b) = 1$ for all $(a, b) \in \Omega$.

%\paragraph{Distance between a word and a regular language}