### 2021-06-03 main-eptcs.tex

% !TEX root = main.tex

% http://www.eptcs.org
\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{GandALF 2021} % Name of the event you are submitting to
\usepackage{breakurl}                 % Not needed if you use pdflatex only.
\usepackage{underscore}               % Only needed if you use pdflatex.

%% theorem environments
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
\def\qed{}
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

\author{Florent Jacquemard
\institute{INRIA \& CNAM, Paris, France\\
\email{florent.jacquemard@inria.fr}}
} % end author

\def\titlerunning{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
\def\authorrunning{Florent Jacquemard}