### 2021-06-10 parse-tree.tex
revise section parsing

Let $\Omega$ be a ranked alphabet
with maximal rank value~$M$, 
and let $\hat\Omega$
be the alphabet with nesting symbols associated to $\Omega$ like in Section~\ref{sec:trees},
for the linearization of trees of $\T(\Omega)$.
%(remember that $\Omegai = \Omega_0$).
%...
%As explained in Section~\ref{sec:trees}, 
%$A = \< Q, \init, \bar{\wei} >$ can be seen as a weighted regular tree grammar, 
%or a labeled weighted CF grammar whose derivation trees are in $\T(\Omega)$, 
%with weight defined by $A$.
%...

we assume in input
(...)
-- a \SWTA $A$ over $\Omega$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: T(\Omega) \to \Semiring$,
(...)

\noindent 
We extend the measure defined by $T$ to 
$d : \Sigma^* \times \T(\Omega) \to \Semiring$ as follows.
Given a word $w \in {\hat\Omega}^*$, the projection of $w$ onto $\Omegai$,
denoted $w|_\Omegai$,
is the word of $\Omegai^*$ obtained from $w$ by removing all symbols 
in $\hat\Omega \setminus \Omegai$.
Using this notation
and the tree linearization operator defined in Section~\ref{sec:trees}, 
$d$ is defined by:
%\begin{equation}
\(
d(s, t) = T\bigl(s, \lin(t)|_\Omegai \bigr)  \mathrm{~for~} s \in \Sigma^*, t\in \T(\Omega)
\).
%\end{equation}

\noindent 
\emph{Symbolic weighted parsing} is the problem, 
given the above input, 
to find a tree $t \in \T(\Omega)$ %nested word $t \in \Omega$ 
minimizing \( d(s, t) \otimes A(t)\)
\wrt $\leq_\oplus$, 
\ie such that: %Hence, it is the problem of finding 
%
\begin{equation}\label{eq:distance-lang}
d(s, t) \otimes A(t) = \displaystyle\bigoplus_{t' \in \T(\Omega)} d(s, t') \otimes A(t') 
\end{equation}
%
The measure expressed in~\eqref{eq:distance-lang} is called
called the edit-distance between~$s$ and~$A$ in~\cite{Mohri03EDWA}.