### 2021-07-20 abstract.tex
passe abstract, introduction

% !TEX root = main.tex
%
% Symbolic Weighted (SW) extension of symbolic automata...
%
We propose a framework for weighted parsing over infinite alphabets.
%
It is based on language models called Symbolic Weighted Automata (\SWA)
at the joint between Symbolic Automata (\SA) and Weighted Automata (\WA),
as well as Transducers (\SWT) and Visibly Pushdown (\SWVPA) variants.
%
Like \SA, \SWA deal with large or infinite input alphabets,
and like \WA, they output a weight value in a semiring domain.
The transitions of \SWA are labeled by functions from an infinite alphabet into the weight domain.
This generalizes \SA, whose transitions are guarded by Boolean predicates
overs symbols in an infinite alphabet,
and also \WA, whose transitions are labeled by constant weight values,
and who deal only with finite alphabets.
%
We present some properties of \SWA, \SWT and \SWVPA models,
that we use to define and solve a variant of parsing
over infinite alphabets.
%
We illustrate the model with a motivating application to
automated music transcription.