### 2021-07-20 main.tex
passe section 3: swT, swA. resubmit.

% SWT without epsilon-transitions
% SWT do not read the other tape in "skip" transitions 10 or 01
%     (hence this version is simpler  than V2)
% SWA without epsilon-transitions
% SW-VPA read top of stack (when not empty) at every transition
%        (6 kinds of transitions).

% default style article.cls
%\input{main-article}

% Springer llncs style
% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
%\input{main-llncs}

% EPTCS style
% http://www.eptcs.org
%\input{main-eptcs}

% LIPICS v.2021 style
% https://submission.dagstuhl.de/documentation/authors
\input{main-lipics}

% comments in margin with \todo command
\usepackage[colorinlistoftodos]{todonotes} % option disable to hide todos
% command todo : options
% - noline: no line connecting the note with the place in the text
%   where the note occurs in the latex code
% - fancyline : curved arrow from note to inssertion point
% - inline: place a todonote inside the text instead of in the margin
% - author=name
\newcommand{\florent}[1]{\todo[noline,size=\tiny,color=yellow!40]{#1}}
\renewcommand{\florent}[1]{} % uncomment for final
\newcommand{\lydia}[1]{\todo[noline,size=\tiny,color=red!30]{#1}}
\renewcommand{\lydia}[1]{} % uncomment for final
\newcommand{\philippe}[1]{\todo[noline,size=\tiny,color=green!30]{#1}}
\renewcommand{\philippe}[1]{} % uncomment for final

\usepackage{hyperref}
%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
%usepackage{url}

% for footnote ref
\usepackage{refcount}

% array and tabular
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

% extension of enumerate env. (style for displaying counters)
% \usepackage{enumerate}

%% pictures
% \usepackage{graphicx}
% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
% \graphicspath{fig/}

%% PGF, Tikz
\usepackage{tikz-cd}
%% \usepackage{pgfplots}
%% \usepgfplotslibrary{dateplot}
%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
% \usepackage{tikz}
% \usetikzlibrary{cd}
%% \usetikzlibrary{arrows}
%% \usetikzlibrary{calc}
%% \usetikzlibrary{snakes}
%% \usetikzlibrary{backgrounds}
% \usetikzlibrary{trees}
%% \usetikzlibrary{automata}
%% \usetikzlibrary{positioning}
%% \usetikzlibrary{matrix}
%% \usetikzlibrary{patterns}
%% \usetikzlibrary{shapes}

% symbols
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{bbold}
\usepackage{latexsym}
%\usepackage{amsfonts}
\usepackage{stmaryrd}
%\usepackage{mathabx}
%\usepackage{MnSymbol}
\usepackage{harmony} % simple music fonts
\usepackage{mathtools} % for arrows
%\usepackage{mathptmx}
\usepackage{comment}

%% algorithms
%\usepackage{algorithm}
%\usepackage{program}
\usepackage[ruled,vlined]{algorithm2e}

%% allows for page break within arrays
\usepackage{longtable}

%% arrows etc
%\input{rewriting}

%% music symbols
% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
\usepackage{musicography}

%% for new macros
\usepackage{xspace}

%% Misc macros
\def\ie{\textit{i.e.}\xspace}
\def\eg{\textit{e.g.}\xspace}
\def\wrt{\textit{wrt}\xspace}
%\def\wlog{\textit{wlog}\xspace}
\def\etc{\textit{etc}\xspace}

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}

%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}}
\newcommand{\config}[2]{\ensuremath{#1}[{#2}]}
\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
\def\stacksep{\cdot}
\def\stackup{\\}

\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}

% \bigominus
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
<-6> mathx5 <6-7> mathx6 <7-8> matha7
<8-9> mathx8 <9-10> mathx9
<10-12> mathx10 <12-> mathx12
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}

\newcommand{\A}{\mathcal{A}} % automata, VPA
\newcommand{\B}{\mathcal{B}} % automata (in constructions)
\newcommand{\C}{\mathcal{C}} % automata (in constructions)
\newcommand{\D}{\mathcal{D}}  % transducer = distance
\newcommand{\E}{\mathbb{E}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}
\newcommand{\seq}{\ensuremath{\mathit{seq}}}
\newcommand{\word}{\mathit{word}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SW{\textsf{sw}\xspace}
\def\swM{\textsf{swM}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\VPA{\textsf{VPA}\xspace}
\def\SVPA{\textsf{sVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\wei{\mathsf{w}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
\def\weiei{\weie[\mathsf{i}]}
\def\weiec{\weie[\mathsf{c}]}
\def\weier{\weie[\mathsf{r}]}
\def\weiex{\weie[\mathsf{x}]}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}
\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
%\newcommand{\ts}[1]{^{#1}} % timestamp marker

\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
\newcommand{\creturn}[1]{\ensuremath \rangle_{#1}}
%\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$

%\sloppy

% Parsing over infinite alphabet as optimal alignment computation
% as edit-distance between string and language
%
%\title{Symbolic Weighted Parsing and Automated Music Transcription}
%\title{Symbolic Weighted Language Models and Automated Music Transcription}
\title{Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets}
%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
%\titlerunning{WVPA \& AMT}
%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
%\authorrunning{Florent Jacquemard}

\date{\today}

\begin{document}
\thispagestyle{empty}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro}
\input{introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% prelin
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminary Notions}
\label{section:prelim}\label{sec:prelim}

%notations: for set $S$ : sets of sequences $S^*$ and $S^+$...
%interval $[i..j]$ of natural numbers

%\paragraph*{Semirings}
\label{section:semiring}\label{sec:semiring}
\input{semiring}

%\paragraph*{Label Theory}
\label{section:symbols}
\input{labels}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SWT & SWA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Automata and Transducers}
\label{section:transducer}\label{sec:transducer}
\label{section:SWA}\label{sec:SWA}
\label{section:SWT}\label{sec:SWT}

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words and languages
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers.
%
The models introduced in this section generalize
weighted automata and transducers~\cite{Droste09handbook}
%over finite alphabets, see \eg~\cite{Mohri03EDWA},
by labeling each transition with a weight function
that takes the input and output symbols as parameters.
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but the latter guards are restricted to the Boolean semiring.
%

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)
\noindent
Let $\Semiring$ be a commutative semiring,
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets},
and $\bar\Phi$ be a label theory over $\Semiring$
containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,
and $\bar{\wei}$ is a triplet of transition functions
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call \emph{number of transitions} of $T$ the number of pairs of states
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions
as functions of
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) &
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}
\]
%
\noindent
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$),
or the opposite ($\wei_{01}$),
or advance in both $s$ and $t$ ($\wei_{11}$).}
% \noindent
$T$ is
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$,
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
% \setminus \{ \< \varepsilon, \varepsilon >\}$,
%
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
 \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}}
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}}
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}}
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}

%
We recall that, by convention (Section~\ref{sec:prelim}),
an empty sum with $\bigoplus$ is equal to~$\zero$.
%
Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
when reading $a$ and $b$ in the input and output words,
increment the current position in the input word if and only if $i = 1$,
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions,
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
symbolic transducers.

%whereas considering the current symbol may be useful to compute a transition weight.
%(even when it does not change the head's position, like with $\varepsilon$-transitions).
%
%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
%are missing in the definition of $\weight_T$.
%It means that $T$ must avoid configurations where it reached the end of
%the output word and not of the input one, or vice-versa.
%This can be done by using $\wei_{10}$ and $\wei_{01}$
%before reaching the end of word, and using a special state for this purpose.

The expression \eqref{eq:SWT-weight}
can be seen as a stateful definition of
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively,
$\wei_{10}(q, a, \varepsilon, r)$ is the cost of
the deletion of the symbol~$a \in \Sigma$ in~$s$,
$\wei_{01}(q, \varepsilon, b, r)$ is the cost
of the insertion of~$b \in \Delta$ in $t$,
and $\wei_{11}(q, a, b, r)$ is the cost
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$,
is the product, with $\otimes$, of the individual costs of the operations involved;
and the distance between $s$ and $t$ is the sum, with $\oplus$,
of all possible products.
%
\medskip\noindent
%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$.
Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is:
%defined as follows:
\begin{equation}
T(s, t)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
\label{eq:SWT-value}
\end{equation}

%\begin{example}
%We can define with a \SWT the computation of a similarity measure
%between timed sequences
%similar to dynamic time warping (DTW).
%
%Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and
%let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
%We consider a \SWT with one state $q$ and transitions
%$\wei_{11}(q, d, d', q) =
% \wei_{01}(q, d, d', q) =
% \wei_{10}(q, d, d', q) = |d' - d|$,
%for all $d, d' \in \mathbb{R}_+$.
%% needs reading input/output symbols even by epsilon-transitions.
%The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
%for the computation of an optimal match between words,
%the matching cost for two symbols being the
%the time distance between them.
%\endex
%\end{example}

\begin{example}\label{ex:SWT}
We build a small \SWT over  alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{}),
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

\begin{comment}
\begin{example}
In Common Western Music Notation~\cite{Gould11Notation},
several symbols may be used to represent one single sounding event.
\lydia{unique $\to$ similar}
\florent{similar $\to$ single}
For instance, several notes can be combined with a tie,
like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}.
\lydia{modif.}
These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
We thus want to be able to compare a music score with music played by a performer.
\lydia{changed end}
%
\noindent
We propose a small weighted transducer model with two states $q_0$ and $q_1$
that calculates the distance between an input sequence of sounding
events (music "performance")
to an output sequence of written events (music "score").
% simple pointwise distance between two sequences of timestamped events **
%
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$
of Figure~\ref{fig:semirings} and
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$
be an output alphabet of symbols with timestamps.
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$,
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let
%$\Sigma = \{ a, - \} \times \R_+$ and
%This example of $\Delta$ is motivated by the case of music notation,
%where several notated events (notes) can be tied together,
%with a \emph{tie} or a \emph{dot}
%meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose
is to compare a recorded performance $s \in \Sigma^*$
with a notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may correspond
to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations
$\<\mathsf{-}, d'_{i+1}>$..., they are just skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also must be able to take performing errors into account, while still being able to compare with the score,
since a performer could, for example, play an unwritten extra note.
\lydia{reformulated this sentence}
%
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$,
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.

\lydia{ccl to the ex}
That way, an \SWT is capable of evaluating the differences between a score and a performance,
all the while ensuring that performance errors are plausible.

%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}
\end{comment}

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based with an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q',\ 
          \mathrm{or}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

%
%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$,
%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

%\subsection{Properties}
\noindent
The following property will be useful for
symbolic weighted parsing (Section~\ref{sec:parsing}).

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
there exists a \SWA $B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
effectively constructible in PTIME, 
such that for all $t \in \Delta^+$,
$B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
\end{proposition}
%
\begin{proof}
(sketch, see Appendix~\ref{sec:closure} for details).
The state set of $B_{A, T}$ is the Cartesian product of
the state sets of $A$ and $T$ and its transitions simulate,
while reading an output word $t \in \Delta^*$, the
synchronized behaviour of $A$ and $T$ on $t$
and some input word $s \in \Sigma^*$.
The weight for reading the input $s$ is obtained with
$\bigoplus^1_\Sigma$.
%
The main difficulty comes from the transitions
of the form $\wei_{10}$, which read in input and ignore the output.
Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
but $\varepsilon$-transitions are not defined for $\SWA$.
Therefore, the $\varepsilon$-transitions
are eliminated on-the-fly during the construction of~$B_{A, T}$,
following a procedure of~\cite{LombardySakarovitch12ciaa}.
%\qed
\end{proof}

%The complexity of construction
%\noindent
%The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
%where the sizes $\| T \|$ and $\| A \|$ are their number of states.
%%of~$T$ is its number of states $|Q|$.
%\florent{revise complexity with nb. of tr. and states? 
%         or skip details and just say polynomial?}
%
The particular case of Proposition~\ref{prop:epsilon} with a singleton $A$,
\ie such that $A(s) = \one$ for a given $s \in \Sigma^*$
and $A(s') = \zero$ for all $s' \neq s$,
corresponds to a construction of a \SWA for the partial application of the \SWT $T$,
fixing the first argument $s$.
\begin{corollary} \label{cor:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and $s \in \Sigma^+$,
there exists an effectively constructible \SWA
$B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
\end{corollary}

%\noindent
%The construction time and size for $B_{T, s}$ are $O(\| T \|^3 . | s |^2)$,
%where the size $\| T \|$ of~$T$ is its number of states $|Q|$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SW-VPA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested word over an infinite alphabet a weight value
in a semiring.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST % parse trees.
(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
%\florent{see \S 5 and App.A}

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Delta$ be a countable alphabet
%finite (large) or infinite,
artitioned into three
subsets~$\Deltai$, $\Deltac$, $\Deltar$,
whose elements are respectively called
\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
%we shall write respectively
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
stand respectively
for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$
%of unary functions in $\Deltai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

\begin{example}\label{ex:nested-word}
In the nested score representation $O \in \Delta^*$ in Ex.~\ref{ex:running},
$\Deltai$ corresponds to timed notes and continuations,
%the set of music events of the form $<\nu, \tau>$
%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
%rest symbol R.
and $\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
%are used to encode the score structure:
%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
%(resp. $\rangle$)
%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
Another example is the other candidate
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
of transcription of~$I$,
linearized into
$O' =$
$\ccall{\mathsf{m}}\ts{[0,1]}$,
$\ccall{2}\ts{[0,1]}$,
$\mbox{\footnotesize A4}\ts{0}$,
$\ccall{2}\ts{[\frac{1}{2},1]}$,
$-\ts{\frac{1}{2}}$,
$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\ccall{\mathsf{m}}\ts{[1,2]}$,
$\ccall{3}\ts{[1,2]}$,
$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
$\mbox{\footnotesize D5}\ts{1}$,
$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
$\creturn{3}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$
(see also Fig.~\ref{fig:score-tree}).
The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
represents an \emph{appogiatura}, %(grace-note),
\ie an ornemental note with theoretical duration~0.
\endex
\end{example}

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states,
$P$ is a finite set of stack symbols,
$\mathsf{in} : Q \to \Semiring$
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state,
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,
$\weiei : Q \times Q \to \Phii$,
$\weic : Q \times P \times Q \times P \to \Phicc$,
$\weiec : Q \times P \times Q \to \Phic$,
$\weir : Q \times P \times Q \to \Phicr$,
$\weier : Q \times Q \to \Phir$.
%and
%$\weiex : Q \times Q \to \Phix$
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
As in Section~\ref{section:transducer},
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$,
$a \in \Deltai$,
$\call{c} \in \Deltac$,
$\return{r} \in \Deltar$,
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Deltai \times Q \to \Semiring &
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Deltar \times Q \to \Semiring &
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}
\]

\noindent
%The intuition is the following for the above transitions.
$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
%\lydia{moved this to the beginning}
%
\noindent
$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
without changing the stack.
Moreover, $\weii$ reads a pair made of
${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
%
\noindent
$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
push it to the stack along with $p'$, and change state from $q$ to to $q'$.
Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
($c$ is compared to $c'$).
%and $\weiec$ applies iff the stack is empty.
%
\noindent
$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
Moreover, $\weir$ reads and pop from stack a pair made of $\call{c}$ and $p$,
($\call{c}$ is compared to $\return{r}$).
%and $\weier$ applies iff the stack is empty.
%In this case, the weight function $\phi_\mathsf{r}$
%computes a value of matching between the call and return symbols $c$ and $r$.
%This value might be set to $\zero$ in order to express that the symbols do not match.

Formally, the transitions of the automaton~$A$ are defined
with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
%
A configuration $q[\gamma]$
is composed of a state $q \in Q$
and a stack content $\gamma \in \Gamma^*$,
where $\Gamma = \Deltac \times P$.
Hence, $\weight_A$ is a function from
$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
The empty stack is denoted by $\bot$, and the topupmost
symbol is the last pushed content.
%
The recursive definition of $\weight_A$
enumerates each of the six possible cases:
reading $a \in \Deltai$,
%\lydia{intro to func}
or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
for each possible state of the stack (empty or not).
%to add to $u \in {\Delta}^*$.
%\lydia{introduced the 6 cases}
\florent{shorter notation $cp$ for $\< c, p>$ ?}

\begin{align}
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero
%\mathrm{~if~} q \neq q'\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u,
               \config{q'}{\gamma'}\bigr) & =
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'')
  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u,
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, a\, u,
               \config{q'}{\gamma'}\bigr) & =
  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'')
   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u,
               \config{q'}{\gamma'}\bigr) & =
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p' \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr)
 \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u,
                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u,
               \config{q'}{\gamma'}\bigr) & =
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p \in P
                          \end{array}}}
  \weiec(q, {\call{c}}, p, q'')
  \otimes \weight_A\bigl(\config{q''}{\<{\call{c}}, p>}, u,
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\return{r}}\, u,
               \config{q'}{\gamma'}\bigr) & =
 {\displaystyle\bigoplus_{q'' \in Q}}
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr)
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u,
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u,
               \config{q'}{\gamma'}\bigr) & =
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'')
  \otimes \weight_A\bigl(\config{q''}{\bot}, u,
                         \config{q'}{\gamma'}\bigr)\nonumber
\end{align}
%\lydia{c p to <c, p>}
%
% and ${\call{c}}\, p\stacksep \gamma$
%denotes a stack where the pair made of ${\call{c}} \in \Deltac$ and $p \in P$ is the top symbol
%and $\gamma$ is the rest of stack.

\noindent
The weight associated by $A$ to $t \in \Delta^*$
is defined according to empty stack semantics:
%
\begin{equation}
A(t)  =
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr)
\mathop{\otimes} \mathsf{out}(q').
\label{eq:SWVPA-value}
\end{equation}

\noindent
Every $\SWA$ $A = \< Q, \init, {\wei_1}, \final >$,
over $\Sigma$, $\Semiring$ and $\bar\Phi$
is a particular case of $\SWVPA$
$\< Q, \emptyset, \init, \bar\wei, \final >$
over $\Delta$, $\Semiring$ and $\bar\Phi$
with $\Deltai = \Sigma$ and $\Deltac = \Deltar = \emptyset$,
and computing with an always empty stack:
$\weiei = \wei_1$ and all the other functions
of~$\bar\wei$ are the constant~$\zero$.

\begin{example} \label{ex:SWVPA}
We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
that expresses a weight related to the music notation,
or more precisely to its structural complexity.
Given a set of equivalent representations, we aim at choosing the simpler one.

%Let us assume a third component in the symbols of $\Deltac$
%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
%quel est "l'associate time interval" d'un symbole. Je ne sais
%pas à quoi fait référence "the following call transition". Je
%n'ose pas trop toucher mais je crois qu'il faut revoir.}
%which is the duration of the associated time interval.
%It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
It reads a new time-division symbol~$\ccall{d}$:
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
The time interval $\iota$ attached to the~$\ccall{d}$ read 
must have a starting time $\tau + \frac{i\ell}{c}$,
where~$\tau$ and~$\ell$ are respectively
the starting time and duration of the interval attached to
the previous time-division symbol read $\ccall{c}$ (found on the stack).
%plus the total duration of intervals read so far, $i \ell$, and
%the new current interval duration is~$\frac{\ell}{d}$.
%
And it must have a duration $\frac{\ell}{d}$.

\noindent
The weight of the above transition is $\alpha_d$.
We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
%
\noindent
Along with $\ccall{d}$,
the above transition pushes
the state $q_{{i+1}/{c}}$ on the stack,
in order to start the next sub-interval after reading 
a return symbol $\creturn{d}$,
with the transition:
\(
\weir\bigl(q_{{d}/{d}},
           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
           \,\creturn{d}\ts{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
\).

\noindent
Reading a musical event $\mu$ is done with:
\(
\weii\bigl(q_{{i}/{d}},
           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{j}/{c}},
           \mu\ts{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
\).
% cas des barres de mesure?
%For the case of a new measure symbol,
%$q_{1/1}$ is a special (initial) state.
The transition to start reading the first measure is:
\(
\weiec\bigl(q_{1/1},
           \ccall{\mathsf{m}}\ts{[0, 1]}, q_{1/1},
           q_{1/1} \bigr) = \one
\),
and for a new measure:
\(
\weic\bigl(q_{1/1},
          \ccall{\mathsf{m}}\ts{[\tau-1, \tau]}, q_{1/1},
          \ccall{\mathsf{m}}\ts{[\tau, \tau+1]}, q_{1/1},
          q_{1/1} \bigr) = \one
\).
\endex
\end{example}

%\subsection{Properties}
\medskip\noindent
Similarly to \VPA~\cite{AlurMadhusudan09nested}
and \SVPA~\cite{dAntonyAlur14SVPDA},
the class of \SWVPA is closed under the binary operators of the underlying semiring.
%
\begin{proposition}\label{prop:SWVPA-product}
Let $A_1$ and $A_2$ be two \SWVPA
over the same $\Delta$, commutative $\Semiring$ and effective $\bar\Phi$.
There exists two effectively constructible $\SWVPA$
$A_1 \oplus A_2$ and $A_1 \otimes A_2$,
such that for all $s \in \Delta^*$,
$(A_1 \oplus A_2)(s) = A_1(s) \oplus A_2(s)$ and
$(A_1 \otimes A_2)(s) = A_1(s) \otimes A_2(s)$.
\end{proposition}
%
\begin{proof}
We do a classical product construction, see Appendix~\ref{sec:product} for details.
%similar to the case of the Boolean semiring~\cite{dAntonyAlur14SVPDA}.
\end{proof}
\noindent
We present now a procedure for searching, for a~$\SWVPA$ $A$,
a word %$w \in \Delta^*$
of minimal weight for~$A$.
%as stated in the following proposition. %\wrt~$\leq_\oplus$,
%It requires the following property of label theories.
%
\begin{proposition}\label{th:best-search}
For a \SWVPA $A$
over $\Delta$,
$\Semiring$ commutative, bounded, total and complete, %semiring
and $\bar\Phi$ effective, %and $k$-convex, % label theory,
one can construct in PTIME a word $t \in \Delta^*$
such that $A(t)$ is minimal \wrt the natural ordering $\leq_\oplus$ for $\Semiring$.
\end{proposition}
%\florent{total?}
%
Let $A = \< Q, P, \init, \bar\wei, \final >$.
%
We propose a Dijkstra algorithm computing,
for every $q, q' \in Q$,
the minimum, \wrt $\leq_\oplus$, of the function
$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
Let us denote by $b_\bot(q, q')$ this minimum.
By definition of $\leq_\oplus$, and since $\Semiring$ is total,
it holds that:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{t\in \Delta^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
\end{equation}
The infinite sum in~\eqref{eq:bbot} is well defined since $\Semiring$ is complete.
%
Following~\eqref{eq:SWVPA-value}, and the associativity, commutativity
and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is:
\begin{equation}\label{eq:min}
{\displaystyle\bigoplus_{t\in \Delta^*}} A(t)
=
{\displaystyle\bigoplus_{t\in \Delta^*}}
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\beta_{q, q'}(t)
\mathop{\otimes} \mathsf{out}(q')
=
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
b_\bot(q, q')
\mathop{\otimes} \mathsf{out}(q')
\end{equation}

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%
Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
mapping every $c \in \Deltac$ to
the minimum weight of a computation of~$A$
starting in state $q$ with a non-empty stack
$\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
and ending in state $q'$ with the same stack $\gamma'$,
such that moreover,
the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
(\ie $\gamma'$ is left untouched during the computation).
However, the computation can read $\<c, p>$ at the top of $\gamma'$,
and can also push another pair $\< c', p'> \in \Gamma$ on $\gamma'$,
following the third case of
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
The pair $\< c', p'>$ can be pop later, during the computation from $q$ to $q'$,
following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
%
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).
%
Formally, in order to define $b_\top$, we consider
a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
representing the above untouched stack, and let:
%
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\quad
\mathrm{~for~all~} c \in \Deltac
\end{equation}
%
By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
using the symbol $\top$ for the part of the stack below $\<c, p>$
(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
ensures that this part is not touched during the computation.
% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
% because it is not possible to push over $\top$ (push read the top of stack).
%
This ensures in particular that the subword
read during the computation is well parenthesized
(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).

% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).

%The term $\config{q}{\bot}, s, \config{q'}{\bot}$
%of this sum is the central expression in
%the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
%of the function $\weight_A$.

\begin{algorithm}
\SetKw{Extract}{extract}
\SetKw{Update}{update}

\textbf{initially} let $\Q = (Q \times Q) \cup (Q \times P \times Q)$,
and let $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \one$
if $q_1 = q_2$ and $d_\bot(q_1, q_2) = d_\top(q_1, p, q_2) = \zero$ otherwise$\;$

\smallskip\noindent
\While{$\Q \neq \emptyset$}{

%\noindent\quad
\Extract $\< q_1, q_2>$ or $\< q_1, p, q_2>$ from $\Q$
%
%\noindent\quad
such that $d_\bot(q_1, q_2)$, 
resp. $\bigoplus_{\Deltac} d_\top(q_1, p, q_2)$, is minimal in $\Semiring$ \wrt $\leq_\oplus$$\;$

%\noindent\quad
%\textbf{for all} $q_0, q_3 \in Q$
\Update $d_\bot$ with $\< q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$ 
(Figure~\ref{fig:best-update}).$\;$
}
\caption{Best search for \SWVPA}
\label{algo:Dijkstra}
\end{algorithm}

Algorithm~\ref{algo:Dijkstra}
constructs iteratively,
using a priority queue $\Q$, 
two markings
$d_\bot : Q \times Q \to \Semiring$ and
$d_\top : Q \times P \times Q \to \Phic$,
%of the triplets $\<q, \sigma, q'>$
%of states of $A$ by weight values in $\Semiring$,
that converges eventually to $b_\top$ and $b_\bot$. 
%\florent{explication Fig.~\ref{fig:best-update} (update) suivant cas de \eqref{eq:SWVPA-weight}?}
%\florent{The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra},
%Figure~\ref{fig:best-update} are well defined since~$\Semiring$ is complete.}
%
% COMPLEXITY
%The algorithm performs $2.|Q|^2$ iterations until $P$ is empty,
%and each iteration has a time complexity $O(|Q|^2 . |P|)$.
%\florent{complexity: detail with and states?}
%That gives a time complexity $O(|Q|^4 . |P|)$.
%It can be reduced by implementing $P$ as a priority queue,
%prioritized by the value returned by $d$.
%** where  $T$ is the worst time complexity for an oracle computing the values
%of Definition~\ref{def:effective}.
%$|Q|^3.\log(|Q|^2)$
%
%Algorithm~\ref{algo:Dijkstra}
It terminates in PTIME and at termination (when $\Q$ is empty), 
and its correctness is ensured by~\eqref{eq:min} and  
the following invariants: 
$\< q_1, q_2> \notin \Q$ iff $d_\bot(q_1, q_2) = b_\bot(q_1, q_2)$, 
and $\< q_1, p, q_2> \notin \Q$ iff 
$\bigoplus_{\Deltac} d_\top(q_1, p, q_2) = \bigoplus_{\Deltac} b_\top(q_1, p, q_2)$.
\florent{(see Appendix~\ref{sec:bestsearch}).}
%of Algorithm~\ref{algo:Dijkstra}.

\begin{figure}[ht]
For all $q_0, q_3 \in Q$, %$p \in P$,
\[
\begin{array}{lcl}
%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
d_\top(q_1, p, q_3) & \opluseq &
  d_\top(q_1, p, q_2) \otimes
  \displaystyle\bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
%
d_\bot(q_1, p, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes
  \displaystyle\bigoplus_{\Deltai} \weiei(q_2, q_3)\\
%
d_\top(q_0, p, q_3) & \opluseq &
  {\displaystyle\bigoplus_{\Deltac}}^2
  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
  d_\top(q_1, p', q_2) \bigr) \otimes_2
  {\displaystyle\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
%
d_\bot(q_0, q_3) & \opluseq &
  {\displaystyle\bigoplus_{\Deltac}}
  \bigl(
  \weiec(q_0, p, q_1) \otimes
   d_\top(q_1, p, q_2) \otimes
  \displaystyle\bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\\
%
d_\bot(q_1, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes
  \displaystyle\bigoplus_{\Deltar} \weier(q_2, q_3)\\
%
d_\top(q_1, p, q_3) & \opluseq &
  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3),
  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
%
d_\bot(q_1, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
\end{array}
\]
%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
%is the partial application
%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
%
%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
%is the partial application
%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
\label{fig:best-update}
\end{figure}
%
%\noindent
%** effectively computable by hypothesis that the label theory is effective**
%\florent{REVOIR}
Thanks to the hypothesis that $\bar\Phi$ is effective, 
it is possible to construct during the iteration a witness word
for Proposition~\ref{th:best-search}, 
\ie a word $t \in \Delta^*$ with a minimal weight $A(t)$ \wrt~$\leq_\oplus$.

%we require the additional property of convexity of weight functions.***
%using the second requirement in Definition~\ref{def:effective}.

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Symbolic Weighted Parsing}
\label{sec:parsing}
%% see parse-tree.tex for a definition with parse tree (instead of nested words)

Let us now apply the models and results of the previous sections %in order to define
to the problem of parsing over an infinite alphabet. %appropriate
%
%\subsection{Definition}
%
Let~$\Sigma$
and~$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$
be countable input and output alphabets,
let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a
commutative, bounded, total and complete %\florent{total?} 
semiring and let $\bar\Phi$ be an effective label theory over $\Semiring$,
containing $\Phi_\Sigma$, $\Phi_{\Sigma, \Deltai}$, as well as
$\Phii$, $\Phic$, $\Phir$, $\Phicr$
(following the notations of Section~\ref{sec:SWVPA-def}).
%
\noindent
We assume given the following input:
\begin{description}
\item[--] a \SWT $T$ over $\Sigma$, $\Deltai$, $\Semiring$, and $\bar\Phi$,
defining a measure %between words
$T: \Sigma^* \times \Deltai^* \to \Semiring$,

\item[--] a \SWVPA $A$ over $\Delta$, $\Semiring$, and $\bar\Phi$,
defining a measure $A: \Delta^* \to \Semiring$,
%a \SWVPA $A$ over $\Delta$, and $\Semiring$, defining a series of nested words
%      $A : \Delta^* \to \Semiring$,
\item[--] an input word $s \in \Sigma^*$.
\end{description}
%
For all $u \in \Sigma^*$ and $t \in \Delta^*$, let
\(d(u, t) = T\bigl(u, t|_\Deltai \bigr)\),
where $t|_\Deltai \in \Deltai^*$ is the projection of $t$ onto~$\Deltai$,
obtained from $t$ by removing all symbols in $\Delta \setminus \Deltai$.
%\end{equation}
%
\noindent
\emph{Symbolic weighted parsing} is the problem,
given the above input,
to find $t \in \Delta^*$ %nested word $t \in \Delta$
minimizing \( d(s, t) \otimes A(t)\)
\wrt $\leq_\oplus$,
\ie s.t. %such that: %Hence, it is the problem of finding
%
\begin{equation}\label{eq:distance-lang}
d(s, t) \otimes A(t) = \displaystyle\bigoplus_{u \in \Delta^*} d(s, u) \otimes A(u)
\end{equation}
%
Following the terminology of~\cite{Mohri03EDWA},
$\SW$-parsing is the problem of computing
the distance~\eqref{eq:distance-lang} between the input~$s$ and the output weighted language of~$A$,
and returning a witness $t$.

\begin{example}[Symbolic Weighted Parsing and the transcription problem]
Applied to the music transcription problem, the above formalism
is interpreted as follows:
%
\begin{itemize}
\item The input word is $I$ of Example~\ref{ex:running}.
\item The \SWT $T$ evaluates a ``fitness measure''  that expresses
    a correspondance between a performance and a nested representation of a music score.
    See Example~\ref{ex:SWT}.
\item The \SWVPA $A$ expresses a cost related to the music notation.
\end{itemize}
%One possibility is to relate this cost to the structural complexity:
%given a set of equivalent representations,
%we aim at choosing the simpler one.
%
As seen in Example~\ref{ex:SWVPA},
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
%\includegraphics[scale=0.20]{pictures/score5.png}
will be favored on
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
%\includegraphics[scale=0.20]{pictures/score4.png}.
when the weight a second time division with $\ccall{2}$ is less
than the difference of weight between
$\mbox{`\footnotesize C$\sharp$5`}$
and $\mbox{\footnotesize C$\sharp$5}$.

\noindent
The SW-parsing framework, applied to the transcription problem,
allows to find an optimal solution
that considers both the fitness of the result,
and its structural complexity.
\endex
\end{example}
The application to music transcription suggested briefly in the examples
has been implemented in a C++ tool~\cite{qparse}, %\footnote{https://qparse.gitlabpages.inria.fr}.
following the principles of the present SW-parsing framework, 
although it differs in several points.
In particular, the automata constructions are performed on the on-the-fly
during the search of a best AST, for efficiency reasons.

%
%\subsection{Computation}
%
\begin{proposition}
The problem of Symbolic Weighted Parsing
can be solved in PTIME in the size of the input \SWT $T$, \SWVPA $A$
and input word $s$,
and the computation time of the functions and operators of the label theory.
\end{proposition}
%
\begin{proof} (sketch)
We follow a \emph{Bar-Hillel} construction for parsing by intersection.
%
\noindent
We first extend the \SWT $T$ over $\Sigma$, $\Deltai$
into a \SWT $T'$ over $\Sigma$ and $\Delta$
(and the same semiring and label theory $\Semiring$ and $\bar\Phi$),
such that for all $u \in \Sigma^*$, and $t \in {\Delta}^*$,
$T'(u, t) = T(u, t|_{\Deltai})$.
%
$T'$ simply skips every symbol
$b \in {\Delta} \setminus \Deltai$
by the addition to $T$,
of new transitions of the form $\wei_{01}(q, \varepsilon, b, q')$.
%
\noindent
Then, using Corolary~\ref{cor:epsilon},
we construct from $s \in \Sigma^*$ and $T'$
a \SWA $B_{s, T'}$,
such that for all $t \in \Delta^*$, $B_{s, T'}(t) = d(s, t)$.
%
%This automaton is such that for all $t \in \Delta^*$,
%\[
%   A_{T', s}\bigl(\lin(t)\bigr)
% = T'\bigl(s, \lin(t)\bigr)
% = T'\bigl(s, \lin(t)|_{\Deltai}\bigr)
% = d(s, t).
%\]
%
\noindent
Next, %we convert the input \SWTA $A$ over $\Delta$
%into a \SWVPA $A'$ over $\hat\Delta$, using Proposition~\ref{lem:SWTA}, and
we compute the \SWVPA $B_{s, T'} \otimes A$,
using Proposition~\ref{prop:SWVPA-product}.
%
\noindent
It remains to compute a best nested word $t \in {\Delta}^*$
using the procedure of Proposition~\ref{th:best-search}.
%and convert it into a best tree in $\T_\Delta$ in order to solve SW parsing
%for~$T$, $A$ and~$s$.
%\qed
\end{proof}

The $\SW$-parsing generalizes
the problem of searching the best derivation (AST) of a weighted CF-grammar $G$
that yields a given input word $w$, called~\emph{weighted parsing},
(see~\cite{Goodman99SemiringParsing}
 and the more general framework~\cite{MorbitzVogler19weighted-parsing}),
with infinite input alphabet instead of a finite one and
transducer-defined distances instead of equality. 
%
See Appendix~\ref{sec:trees} for more details on the correspondence between 
nested words $t \in \Delta^*$ and AST and CF grammars and \SWVPA.

%
%Indeed, it corresponds to the case where $T$
%accepts only the pairs $\<s, t>$ such that
%$s$ is the projection of $t$ on $\Deltai$.
%This can be done with a single state $q$ and
%with transition rules of the form:
%\begin{description}
%\item[] $\wei(q, \varepsilon, a, q) = \one$ for all $a \in \Deltac \cup \Deltar$,
%\item[] $\wei(q, a, a, q) = \one$ for all $a \in \Deltai$,
%\item[] $\wei(q, a, b, q) = \zero$ for all $a, b \in \Deltai$, $a \neq b$.
%\end{description}
%
%Weighted parsing
%corresponds to $\SW$-parsing in the case of finite alphabets,
%a transducer $T$ computing the identity and some \SWVPA~$A$
%obtained from $G$. %the weighted CF grammar.
%Indeed, the \emph{depth-first} traversal of an AST $\tau$
%yields a well-parenthesised word $\lin(\tau)$ over an alphabet
%$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$,
%assuming \eg that $\Deltai$ contains the symbols labelling the leaves of $\tau$ 
%(symbols of rank $0$),
%and $\Deltac$ and $\Deltar$ contain respectively one left and right parenthesis
%$\ccall{b}$ and $\creturn{b}$ for each symbol $b$ labelling inner nodes of $\tau$ 
%(symbols of rank $>0$).
%
%With this representation, the projection $\lin(t)|_\Deltai$ is the
%sequence of leaves of $\tau$. %, enumerated in a \emph{dfs}-traversal.
%
%We show in Appendix~\ref{sec:trees} how to construct
%%($\SW$) tree automaton~$A$ %(in particular the TA computing on the AST of a given CF-grammar)
%a $\SWVPA$ $A$ such that $A(\lin(\tau))$ is the weight the AST $\tau$ of $G$.

%That also holds for the set of ASTs of a weighted CF-grammar.
%The purpose of the transducer $T$ is to measure a distance between input and output words,
%\ie between the word $s$ and the sequence leaves of a derivation tree, in $\Deltai^*$.
%
%that generates (weighted) trees by replacement of a state symbol~$q_0$ (non-terminal),
%by a tree $a(q_1,\ldots, q_k)$, where $k = \rank(a)$.
%A replacement rule $q_0 \to a(q_1,\ldots, q_k)$,
%of weight $\wei(q_0, a, q_1 \ldots q_k) \in \Semiring$
%according to Definition~\ref{def:SWTA},
%corresponds to the production rule $q_0 := a(q_1,\ldots, q_k)$ of a weighted CF grammar,
%with set non-terminal symbols $Q$ and set of terminal symbols $\Delta_0$.
%
%This actually is a slight generalization of CFG since
%each such production rule is labelled by  a symbol of $\Delta_{>0}$,
%hence parse trees %derivation trees
%are trees of $\T_\Delta)$.
%
%Another (more original) generalization is that the set of terminal symbols
%$\Delta_0$ may be infinite.
%
%
%The input language can also be expressed as a \SWTA, or,
%as a particular case, as a weighted context-free grammar,
%converted in turn into a \SWVPA following Lemma~\ref{lem:SWTA}.
%
%\florent{In practice,
%the functions in transition might be represented by diagram such as
%Algebraic Decision Diagrams~\cite{Bahar97ADD}.}
%\paragraph{Application to Automated Music Transcription.}
%\input{transcription}
%...

\section*{Conclusion}

We  introduced Symbolic Weighted language models %over infinite alphabets,
and applied them to the problem of parsing
with infinitely many possible input symbols (typically timed events).
Our approach extends conventional parsing and weighted parsing
by computing a derivation tree modulo
a distance between words
defined by a SW transducer given in input.
This allows to consider finer word relationships than strict equality.
%

This work can be extended in several directions. First, the
the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). Second,
%we believe that
the complexity bounds of the algorithms could be more precisely characterized,
as well as expressiveness of \swM compared to the automata of 
\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
Finally, the  best search algorithm presented here offline, whereas
an on-the-fly automata construction would allow for online parsing,
a suitable feature in the context of applications such as, 
\eg automatic music transcription.

\florent{comparison of expressiveness with the automata 
of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BIBLIO                                                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{plain}
%\bibliographystyle{abbrv}
\bibliographystyle{plainurl}  % LIPICS
%\bibliographystyle{splncs04} % LLNCS
%\bibliographystyle{eptcs}    % EPTCS
%\bibliography{generic}

\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% APPENDIX                                                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix

\section{Properties of Label Theory Operators} \label{sec:closure}
\input{app-labels}

\section{Proof of Proposition~\ref{prop:epsilon}} \label{sec:closure}
\input{app-closure}

\section{Proof of Proposition~\ref{prop:SWVPA-product}} \label{sec:product}
\input{app-product}

%\section{Correctness of the Best-Search Algorithm} \label{sec:bestsearch}
%\input{app-bestsearch}

\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
\input{app-trees}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TODO                                                                       %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\listoftodos

\end{document}