### 2021-09-01 FST-TCS/rebuttal/response.txt
draft response

We would like to thank the reviewers for their valuable work, 
and their comments and references/suggestion.

In particular, we are grateful to them for pointing us to the work of L. Herrmann & Heiko... in DLT 2016 & PhD 2020.  
Indeed, the models that we are using in the paper are particular (restricted) cases of the Weighted Symbolic Automata with Storage defined in DLT 2016 & PhD 2020.
We agree that it should definitely be explained at the beginning of the paper, and the 2 references must be cited as well.

However, It seems difficult to completely skip definitions of the models, for self-completeness reasons.
definition of models less general sufficient for our purpose
not clear whether Prop. 10 (in center pages) is a direct consequence of Hermann V. 2016...

also agree to focus the paper on its 2 main contributions, which are the following:
- a best-search algorithm
  (as far as we know this is a new result
  no such algorithm has been proposed for swVPA)
- a framework for parsing over infinite alphabet and with tree-automata. 
  in particular the use of weighted transducers for def. of distance and of VPA for computing on linearization of parse trees.

second is new application 
Moreover, 
Although this only mentioned quickly in the paper, our main motivation / purpose is a formal study of an empirical (challenging) problem : Automated Music Transcription, based on consistent theoretical framework made of formal language models / formalisms.
in parallel to developpement of a transcription tool
so far results have shown that theoretical foundations improve accuracy…

Regarding the best-search algorithm, 
answer to reviewer 2:... (DETAILS)

questions of reviewer 3 about the distance and the work of Mohri.

About the weighted parsing : rem. reviewer1 on "previous approaches" which are more general.

other rem. review 1:
the relationship between VPA and TA has been discussed in several places, 
**we did [or not?]** find an explicit construction of VPA from TA.
could be move to paper if you think that it is original.
[ ] check in paper [5]