### 2021-10-11 proof-best.tex
include proofs in text. rename files.

\textbf{Proof of correctness}

To prove that the algorithm does compute the word with the best weight, we use the following invariant :
$$
\<q_1, q_2> \notin Q \iff d_{\bot}(q_1, q_2) = b_{\bot}(q_1, q_2)
$$
In other words, a tuple $\<q_1, q_2>$ is removed from the queue during an iteration only when $d_{\bot}(q1, q2)$ has its minimal value.

Initialization : Before the while loop, no tuple has been extracted from the queue, so the invariant is trivially true.

Recursion :

Let's show by contradiction that at each iteration of the while loop, for the tuple $\<q, q'>$ extracted from Q we have:
$$
d_{\bot}(q, q') = b_{\bot}(q, q'), 
$$

Let's consider the states $q_0$, $q_1$, $q_2$ and $q_3$ $\in Q$ such as they form a path $q$ ~...~>$q_0$ -> $q_1$ -> $q_2$ -> $q_3$ ~..~> $q'$.

We take $\<q_0, q_1>$ as the first tuple to be removed from the queue. Then when $\<q_0, q_1>$ is extracted from the queue, the algorithms updates $d_{\bot}(q_0, q_2)$ and $d_{\bot}(q_0, q_3)$. Thus, we have:
$$
d_{\bot}(q_0, q_2) \geq b_{\bot}(q_0, q_1) \\
d_{\bot}(q_0, q_3) \geq b_{\bot}(q_0, q_1)
$$
Let's show that it is true for each update possible, shown in Figure 3. TODO

Let $\<q_0, q_3>$ be the first tuple for which the invariant is false:
$$
For \<q_0, q_3> \notin Q, d_{\bot} (q_0, q_3) \neq b_{\bot} (q_0, q_3)
$$
Since $d_{\bot}(q_0, q_2)$ has to be computed to compute $d_{\bot}(q_0, q3)$, and because the invariant is true for  $d_{\bot}(q_0, q_2)$ we can say:
$$
d_{\bot}(q_0, q_2) = b_{\bot}(q_0, q_2) \\
d_{\bot}(q_0, q_2) \leq b_{\bot}(q_0, q_3) \\
\\d_{\bot}(q_0, q_2) \leq d_{\bot}(q_0, q_3) \\
$$
 However, when $\<q_0, q_3>$ was extracted, $\<q_0, q_2>$ was still in $Q$. Thus, by definition 
$$
d_{\bot}(q_0, q_2) \geq d_{\bot}(q_0, q_3) \\
$$
Then the two inequalities are actually equal : 
$$
d_{\bot}(q_0, q_2) = b_{\bot}(q_0, q_2) = d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3) \\
$$
We find that $d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{\bot}(q_0, q_3) = b_{\bot}(q_0, q_3)$ iff 
$\<q_0, q_3> \notin Q$.

Termination: TODO